S-Aligner: Ultra-scalable read mapping on Sunway Taihu Light

### 2017-04-05 s-aligner.tex
cleanup of content
 - remove executable permissions
 - remove backup files
 - move files to directories

\documentclass{sig-alternate-05-2015}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage{cases}

% REMOVE ME: added for revision
% \usepackage{color}

% for the code listing
\usepackage{listings}
\lstset{language=C++,
    basicstyle=\ttfamily \footnotesize,
    keywordstyle=\color{black}\ttfamily,
    stringstyle=\color{black}\ttfamily,
    commentstyle=\color{gray}\ttfamily,
    morecomment=[s][\color{blue}]{/**}{**/},
    showstringspaces=false,
    showtabs=false,
    captionpos=b,
    numbers=left
}

@@

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\doi{10.475/123_4}

% ISBN
%\isbn{123-4567-24-567/08/06}

%Conference
%\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

%\acmPrice{\$15.00}

%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{S-Aligner: Ultra-scalable read mapping on Sunway Taihu Light}
%\subtitle{[Extended Abstract]}
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}

\numberofauthors{7} 
%Xiaohui Duan, Kai Xu, Yuandong Chan, Christian Hundt,\\ Pavan Balaji, Bertil Schmidt, Weiguo Liu\\
\author{
\alignauthor 
%Xiaohui Duan\\
%    \affaddr{Shandong University}\\
%    \affaddr{Jinan City, China}\\
%    \email{sunrise.duan@mail.\\sdu.edu.cn}\\
\alignauthor
%Kai Xu\titlenote{The first two authors contribute equally to the paper.}\\
%    \affaddr{Shandong University}\\
%    \affaddr{Jinan City, China}\\
%    \email{xu\_kai@mail.sdu.edu.cn}\\
\alignauthor
%Yuandong Chan\\
%    \affaddr{Shandong University}\\
%    \affaddr{Jinan City, China}\\
%    \email{yuandong.c@mail.sdu.edu.cn}\\
\and
\alignauthor
%Christian Hundt\\
%    \affaddr{Johannes Gutenberg University}\\
%    \affaddr{Mainz, Germany}\\
%    \email{hundt@uni-mainz.de}\\
\alignauthor
%Pavan Balaji\\
%    \affaddr{Argonne National Laboratory}
%    \affaddr{Argonne, Illinois, USA}\\
%    \email{balaji@anl.gov}\\
\alignauthor
%Bertil Schmidt\\
%    \affaddr{Johannes Gutenberg University}\\
%    \affaddr{Mainz, Germany}\\
%    \email{bertil.schmidt@uni-mainz.de}\\
\and
\alignauthor
%Weiguo Liu\\
%    \affaddr{Shandong University}\\
%    \affaddr{Jinan City, China}\\
%    \email{weiguo.liu@sdu.edu.cn}\\
}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=\linewidth]{SW26010}
		\caption{Architecture of the SW26010 processor, which is partitioned into four CGs.}
		\label{fig:SW26010}
	\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\includegraphics[width=1\linewidth]{BitDP}
	\caption{Relations between the original matrix, $\varDelta v$, $V^+$, and $V^-$.  Here (a) shows the original score matrix; light-shaded cells indicate $\varDelta v_{i,j}=+1$ while heavily shaded cells correspond to $\varDelta v_{i,j}=-1$. The vertical derivative $\varDelta v$ shown in (b) is subsequently one-hot encoded by the bit vectors $V^+$ and $V^-$ in (c) and (d), respectively.}
	\label{BitDP}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{GridNew}
	\caption{Our task-partitioning and file-loading strategy: (a) overall design; (b) and (c) detailed behavior of processes.}
	\label{TaskGrid}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=0.9\linewidth]{RefDiv}
		\caption{Exemplary partitioning of a reference genome based on concatenation and padded splits.}
		\label{RefDiv}
	\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{AsyncRead}
	\caption{Asynchronous data loading strategy: (a) default case, and (b) case when there is no significantly shorter reference block.}
	\label{AsyncRead}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{FrmWk}
	\caption{Workflow of S-Aligner on a single CG: gray boxes correspond to tasks assigned to SPs while tasks in white boxes are executed on the MP.}
	\label{FrmWk}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{AsyncPara}
	\caption{Asynchronous filtration.}
	\label{AsyncPara}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{SPParallel}
	\caption{Framework of our parallelization scheme within one CG: the MP performs filtration and output of alignment results while SPs verify identified candidate locations and generate base-pair level alignments. Gray boxes indicate the intervals verified in the second round of verification (assuming $slice\_num = 4$).}
	\label{SPParallel}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=0.9\linewidth]{MixPack}
	\caption{Bit-level encoding strategies for DNA sequences. For simplicity, we use a 4-bit word instead of 64-bit word for illustrating our combined AoSoA strategy (bottom).}
	\label{MixPack}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{AsyncTrans}
	\caption{Asynchronous data transfer to LDM.}
	\label{AsyncTrans}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{VarVerCha}
		\caption{Comparison of the runtimes of four implementations of Myers' algorithm on a single CG.}
		\label{VarVerCha}
	\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{VarOptCha}
		\caption{Evaluation with and without I/O optimizations.}
		\label{VarOptCha}
	\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{VarProcCha}
		\caption{Speedup and efficiency of S-Aligner for different numbers of nodes.}
		\label{VarProcCha}
	\end{center}
\end{figure}

@@

%replace this by a "lessons learned" paragraph for Sunway Taihu Light programming
%Possible directions of future research are 1\textsuperscript{st} the design of a dynamic scheduler based on MPI's one-sided communication primitives to allow for a better load balancing, and 2\textsuperscript{nd} the porting of MP-exclusive tasks to SPs in order to further increase performance.

\bibliographystyle{plain}
\bibliography{references.bib}
% That's all folks!
\end{document}

%I'm very tired to navigate over the abandon paragraphs in the tex, thus I moved them here or in the version of Dec 18, 10:37pm.

%Formerly in Abstract:
%About the file system:
%Moreover, since each node has a low file system bandwidth but and the file system will have performance down if there are too many requests, so asynchronous access and data sharing strategy are employed during file IO to maximize the bandwidth to the file system.	{\color{red} Xiaohui, further specify the aforementioned sentences, please. You know the bottlenecks best!}

%About arch:

% REMARK (Christian): We should move this to the introduction
%The Sunway Taihu Light supercomputer has 40,960 nodes, each node has four computation groups (CG), each of CG has a master processor (MP) and a slave processor (SP) group formed by 64 slave processors, and supports ShenWei's SIMD extension, so we employ a flexible data partition strategy among CGs and a two-level parallelization  strategy in a single CG according to Sunway's architecture, and achieve an efficient utilization of the entire system.

%In our implementation, we use MP for lookup works and SP for computation works. The most compute-bounded stage of S-Aligner is the verification of large amounts of potential mapping intervals by Myers algorithm, and we accelerate this stage by using the SP group. Since the memory bandwidth is shared by all processors in one CG, we also use SPs to parallelize a large amount of discrete memory access, also, asynchronous IO strategies are applied to reduce the latency of data transfer.

%Just rewrite:

%Since our algorithm is data-intensive as well, we divide MPI processes to groups with hundreds of processes, and use task grid and group-and-broadcast strategy to transfer shared data, and with the number of nodes grows, our implementation gains almost linear speedup on up to 13,312 nodes with 53,428 processes.

%Formerly in Introduction:

%With the rapid advance of NGS technology, computational biology has experienced great development, and the NGS technology are now widely applied to various fields including CHIP-seq, RNA expression profiling through RNA-seq, DNA Structural Variation detection, and Genome Assembly.

%The emergence of NGS instruments, such as Pacific Biosciences and Illumina MiSeq sequencers\cite{reviewngs}, can produce billions of short sequence data (reads) with length from 25 to 15,000 base pairs (bp) each in a day.

%To make use of these reads, the most important and time-consuming step is to map them to a known DNA sequence (reference), generating bp-level alignments between each read and its potential originating region on the reference. Alignment are required by most DNA sequence analysis, such as genotyping, SNP discovery and personal genomics.

%References contain a mass of base-pairs, for example, there are more than 3 billions of bps in one human genome reference, but traditional local comparison algorithms like Smith-Waterman\cite{sw} usually come with quadratic time complexity computing the similarity between a read and all substrings from a reference. Thus applying this kind of algorithms to read mapping can take  several weeks even months if serially processed on a personal computer.

%There are two modes of read mapping: "any-best" mode and "all-best" mode\cite{allany}, for example, Bowtie\cite{bowtie}, Bowtie2\cite{bowtie2}, BWA\cite{bwa} and GEM\cite{gem} run in "any-best" mode, and they try to identify one or a few best mapping intervals for each read, while RazerS3\cite{razers3}, Hobbes2\cite{hobbes2}, BitMapper\cite{bitmapper} and mrFAST\cite{mrfast} run in "all-best" mode, that is, they find all mapping intervals satisfying some constraint (eg. under a restricted edit distance).

%Obviously "any-best" mappers are usually faster than "all-best" mappers, but some researches need the information about all potential originating intervals of a read from a reference genome, for example, it is much easier to study gene innovation and phenotypic variation if a mapping result of an "all-best" mapper is provided. Therefore S-Aligner is developed as an "all-best" mapper, fulfilling requirements of different applications, and it comes with good scalability, so it can be faster than many "any-best" mappers if enough nodes are provided.
%With the explosive growth of sequence data size, some recent sequencing data can make mapper take several days on a workstation-level computer to compute the alignment, so mapping enormous reads has become a challenge in computational biology. Thus, it is essential and practical to port alignment tools to clusters or supercomputers.
%Supercomputers often comes with co-processors like GPU, Xeon Phi or customized ones, and these co-processors often come with significant performance improvement and better energy/cost efficiency in scientific computing. But how to make full use of these co-processors in computational biology becomes a problem.
%Recently released supercomputer Sunway Taihu Light comes with an architecture  like a neo-heterogeneous platform somehow. The Sunway Taihu Light contains 40,960 nodes equipped SW26010 processor, with 163,840 CGs in total, and each CG has one MP and 64 SPs, while both MP and SP from Sunway Taihu Light are compatible to ShenWei's RISC instruction set except some extensions are not available on SP, so it is regarded as a good platform for sequence alignment problem because both MP and SP have a good performance to handle both scalar and vector operations.

%In this paper, we present our work of a fast and efficient read mapper on SW26010 processor with MPI capability, and it has been evaluated on Sunway Taihu Light since biology computing problem is often popular on most highly ranked supercomputers. We introduce an efficient implementation to find all possible intervals within a reference sequence that matches a read with restricted edit distance. In summary, the main contribution of our work are:

%\begin{enumerate}
% 	\item
% 	In the inter CG level, we use MPI technology to make full use of multiple CGs and apply task grid and broadcast in row/column strategy to achieve a balanced workload and efficient parallelization.
% 	\item
% 	In the intra CG level, we use task parallelism in one CG: using MP for reference index lookup and misc operations like IO or data transfer and using the SP group for the most compute-bounded verification and alignment operation, and we use data parallelism multi-threading among SPs.
% 	\item
% 	In the intra SP level, we use ShenWei's SIMD extension to accelerate the Myers algorithm.
% 	\item
% 	Besides, we use many asynchronous strategy to reduce IO overhead, including file reading among CGs, computing between MP and the SP group and task data fetching on SPs, which greatly decreased processors' idle time.
    %\item
    %We implement the Myers' algorithm using SPs and parallelize it using athreads, which is thousands times faster than the single thread naive implementation of edit distance.
	%\item
	%We make use of LDM via DMA-intrinsics, which gains 22 times speed up over the non-DMA implementation.
	%\item
	%We employ efficient inter CG parallelization strategy, and it can achieve an efficiency of over 97\% with 53,248 CGs.
%\end{enumerate}

%\begin{itemize}
%	\item
%	In Section \ref{Related Work}, we introduce some related algorithms, implementations, and the Sunway Taihu Light supercomputer.
%	\item
%	In Section \ref{Implementation}, we present the implementation in detail, including intra SP, intra CG and inter CG optimization.
%	\item
%	In Section \ref{Evaluation}, we show our evaluation results, including single node and mass parallel experiments.
%	\item
%	In Section \ref{Conclusion}, we concludes the paper.
%\end{itemize}

%Formerly in Related work:

%{\color{green} try write something about razers3}

%corresponding intrinsics library is provided by \textit{ShenWei 5 Compiler Collection} (\textit{sw5cc}), and \textit{sw5cc} also contains an MPI runtime compatible with \textit{mpich2-3.04}.

%For supporting SPs, ShenWei provides an \textit{athread} library, which can start threads on SPs, there are mainly two method to start threads in \textit{athread} library, one is \textit{create}, which creates one thread with an optional parameter and bind it to one SP, the other is \textit{spawn}, which creates a thread for each SP, while these threads share an optional parameter.

%\subsection{Pigeonhole Principle}
%{\color{red}
%The number of potentially matching positions in the reference genome can be drastically reduced by applying filters to the candidate set from the seed stage. S-Aligner employs such a preprocessing filter based on the \emph{pigeonhole principle}. To put it in simple words, if $n$ items are put into $n+1$ boxes that one or more boxes must be empty.
%%For sequence alignment problem, we note the reference genome as $ref$ and the read sequence as $read$, and $s[i, j]$ as the substring of string $s$ starting from $i$ with length $j$, then we define $q\_gram$ and $L_q$ as follows:

%In the context of sequence alignment, let us refer to the reference genome as $ref$ and to the read as $read$. Moreover, the contiguous subsequence of the string $s$ starting at position $i$ with length $j$ shall be denoted as $s[i, j]$. Correspondingly, a $q$-gram is defined as strided substring of length $q$: $q\text{-gram}(i) := read[i \cdot q, q]$. Furthermore, let
%%\begin{equation}
% \label{LQ}
% L_q(i):=\{x | ref[x + i \cdot q, q] = q\_gram(i)\} \quad.
% \end{equation}

% Then we divide read into $n=\lfloor \cfrac{|read|}{q}\rfloor$ $q\_gram$s, from $q\_gram(0)$ to $q\_gram(n - 1)$ and $L_q(i)$ is the set of intervals $l$ that if $read$ is mapped to $l$th bp of $ref$, $q\_gram(i)$ will be exactly mapped, additionally, we define $M_e$ is the set of intervals that $read$ can be mapped to with edit distance at most $e$.

% Then if a read can be mapped to a interval on the reference with at most $e$ edits, at least $n - e$ of the non-overlapping q-grams can be mapped to the reference exactly because there are at most $e$ edits, so if $n > e$, we have
% \begin{equation}
% M_e \subseteq \bigcup^{n}_{i=0}L_q(i)
% \end{equation}.

% Moreover, we have
% \begin{equation}
% \label{MeinG}
% M_e \subseteq \bigcup_{i \in G}L_q(i)
% \end{equation}
% for any $G$ that $G \subseteq [0, n) \land |G| > e$ because at least one q-gram can be mapped exactly.

% For reducing the number intervals for further filtration, we may choose the $G$ having minimum
% \begin{equation}
% \sum_{i \in G}|L_q(i)|
% \end{equation}
% . According to (\ref{MeinG}), this can be done by selecting $e + 1$ $L_q$s with minimum $|L_q|$.

% THE WHOLE SUBSECTION IS BARELY READABLE AND HAS TO BE REVISED THOROUGHLY: BETTER EXPLAIN THE SETS M\_e and L\_q !  
%}
%{\color{green} Weiguo suggests to move this to algorithms section}

%Edit distance is one of the most popular measures for interval verification with restricted number of edit operations $e$. Similar to Smith-Waterman's algorithm, edit distance is computationally demanding since its dynamic programming algorithm exhibits a time complexity proportional to the product of sequence lengths. Let $n$ and $m$ be the lengths of two compared strings $s$ and $s'$ then their edit distance can be determined by relaxing the cells of the $(n+1) \times (m+1)$ shaped cost matrix $C$ according to the recursive scheme
%\begin{equation}
%\delta_{ij}=\left\{
%\begin{array}{rcl}
%0,     &      & {ref\_map_{i}=read_{j}},\\
%1,     &      & {otherwise}\\
%\end{array} \right.
%\end{equation} 
% where $0 < i \le m$, $0 < j \le n$, and the characteristic function $\chi(x \neq y)$ being one for all $x \neq y$ and zero otherwise. 
%It is obviously the time complexity of calculating this DP matrix is $O(n\cdot m)$, as $m$ is often close to $n$, the complexity is also regarded as $O(n^2)$. It is slow to calculate the matrix with a naive recurrence method, but 
% Myers proposed a bit-vector algorithm \cite{myers} that exploits bit parallelism by encoding states in an integer representation. To achieve that the differences (deltas) between adjacent rows and columns in the cost matrix $C$ are used instead of their absolute values. The adjacent encoding reads as follows
% \begin{alignat}{100}
% \varDelta v_{i,j}&=C[i,j]-C[i-1,j]   &&\in \{0, \pm1\} \nonumber\\
% \varDelta h_{i,j}&=C[i,j]-C[i,j-1]   &&\in \{0, \pm1\} \nonumber\\
% \varDelta d_{i,j}&=C[i,j]-C[i-1,j-1] &&\in \{0, +1\}   \quad,
% \end{alignat}
% where $\varDelta h_{i,j}$,  $\varDelta v_{i,j}$ and  $\varDelta d_{i,j}$ are the discrete derivatives of $C[i, j]$ in direction of its left, upper, and diagonal cell. Note that their absolute value is either zero or one due to the uniform choice of gap penalties and substitution costs in Equation~\ref{update_scheme}. As a result, we can encode the relatively small state space with the help of five vectors:
%and the absolute value of them is never greater than 1, and it is proved in \cite{myers}, so we can encode $\varDelta h_{i,j}$,  $\varDelta v_{i,j}$ and  $\varDelta d_{i,j}$ to following bit-vectors:
% \begin{alignat}{100}
% V^\pm_{i,j} &= \chi(\varDelta v_{i, j} = \pm1) \nonumber\\
% H^\pm_{i,j} &= \chi(\varDelta h_{i, j} = \pm1) \nonumber\\
% D^0_{i,j}   &= \chi(\varDelta d_{i, j} = \phantom{\pm}0) 
% \label{one-hot-myers}
% \end{alignat}
%{\color{green} what does $\chi(s_{i-1} \neq s'_{j-1})$ mean, if we can write formulas above like this:
%$$
%V^\pm_{i,j} = \chi(\varDelta v_{i, j} = \pm1)
%$$
%}
%Figure~\ref{BitDP} shows an example for the encoding of the vertical derivative $\varDelta v$ and its associated one-hot representations $V^+$ and $V^-$. Using the bit-vector representation from Equation~\ref{one-hot-myers} the cell updates can be rewritten as follows \cite{myers}:

% \begin{alignat}{100}
% H^-_{i, j} &= V^+_{i, j-1} &&\land D^0_{i, j} \nonumber\\ 
% V^-_{i, j} &= H^+_{i-1, j} &&\land D^0_{i, j} \nonumber\\
% H^+_{i, j} &= V^-_{i, j-1} &&\lor \lnot(V^+_{i, j-1} \,\lor D^0_{i, j})   \nonumber\\
% V^+_{i, j} &= H^-_{i-1, j} &&\lor \lnot(H^+_{i-1, j}   \lor D^0_{i, j})   \nonumber\\
% D^0_{i,j}  &= V^-_{i, j-1} &&\lor H^-_{i-1, j} \lor \chi(s_{i-1} = s'_{j-1})\quad. \label{EVform}
% \end{alignat}

%Formerly in Implementation:

%Afterwards, we hash all 12-mers from the reference to a 24 bit collision-free hash map. The keys are implicitly defined by the slot indices and the values correspond to the interval positions in the reference. Figure~\ref{HIdx} shows the memory organization of the buckets. Multiple values are simply stored in a contiguous 1-d array $hv\_pos$, while another array $hv\_index$ stores the offsets to the buckets. This can be achieved in linear time using default array compaction techniques based on adjacent comparisons, a subsequent prefix scan, and a final substitution phase.

%The resulting index is relatively large. Let $n$ be the length of a chromosome then its concatenated DNA sequence has a length of $2n$. The size of $hv\_pos$ corresponds to $2n\log(2n)$ bits since we need at least $\log(n)$ bits to represent the index space. Analogous, the array of offsets $hv\_index$ occupies $2^{24}\log(2n)$ bits. Note that variable word length may hurt performance and thus we stick to 32 bit integers which is enough for the length of a chromosome.

%This results in an overall size of $(32 \cdot 2n + 32 \cdot 2^{24}) \text{ bits} = (8n+64M) \text{ bytes}$. Thus, for a chromosome of size 300 MB the index occupies roughly 2,464 MB which is more than one-third of the available memory of a CG. Nevertheless, the described data structure exhibits reasonable performance due to its cache-friendly memory layout.

%Since the alignment generation is less often called than the verification, the efficiency of this stage is less important, we implemented a naive Smith-Waterman algorithm to generate alignments on the SPs with only one optimization: for we just fill cell $(i, j)$ if $|i - j| \leq e$ (where $e$ is the edit distance restriction defined in last subsection), and the time taken by alignment in this step is about 1/1000 of the verification, so this step is less worthy to optimize.

%\subsection{Data Structures}
%\label{DataStructure}

%\subsubsection{Reference Index}
%{\color{green} Weiguo suggests it should be removed}
% We construct a hash index for each reference block mentioned in Section \ref{DataStructure}. In particular, we replace 'N's in the block with random base pairs, and subsequently append the block's reverse complement to the end of this block. As a result, we can stick to the 2 bit encoding of nucleotides and do not have to waste an additional bit for 'N'. Moreover, we can ignore reversed reads during mapping.
% %so that we can use 2 bits to encode a bps and we do not need to reverse reads while computing mapping, and we note the modified sequence as $ref$.

% %Then we hash all 12-mers from the $ref$ to a 24-bit collision-free hash value, and store each k-mer's index on the $ref$ to a bucket corresponding to its hash value, Fig.\ref{HIdx} shows our memory organization of these buckets, we use 1-d array $hv\_pos$ to store elements of the buckets, while another array $hv\_index$ stores the start index of buckets.
% Afterwards, we hash all 12-mers from the reference to a 24 bit collision-free hash map. The keys are implicitly defined by the slot indices and the values correspond to the interval positions in the reference. Figure~\ref{HIdx} shows the memory organization of the buckets. Multiple values are simply stored in a contiguous 1-d array $hv\_pos$, while another array $hv\_index$ stores the offsets to the buckets. This can be achieved in linear time using default array compaction techniques based on adjacent comparisons, a subsequent prefix scan, and a final substitution phase.

% %This can be done by a 2-round scanning of $ref$. In the first round, we count the numbers of k-mers grouping by their hash value, and store them in $hv\_index$, then we do a accumulate sum of $hv\_index$ and determine the index to put elements. In the second round, we put the index of each k-mer on the $ref$ to corresponding position of $hv\_pos$.

% The resulting index is relatively large. Let $n$ be the length of a chromosome then its concatenated DNA sequence has a length of $2n$. The size of $hv\_pos$ corresponds to $2n\log(2n)$ bits since we need at least $\log(n)$ bits to represent the index space. Analogous, the array of offsets $hv\_index$ occupies $2^{24}\log(2n)$ bits. Note that variable word length may hurt performance and thus we stick to 32 bit integers which is enough for the length of a chromosome.
% %, we use 32 bit integer type to store $hv\_pos$ and $hv\_index$,
% This results in an overall size of $(32 \cdot 2n + 32 \cdot 2^{24}) \text{ bits} = (8n+64M) \text{ bytes}$. Thus, for a chromosome of size 300 MB the index occupies roughly 2,464 MB which is more than one-third of the available memory of a CG. Nevertheless, the described data structure exhibits reasonable performance due to its cache-friendly memory layout.

% %But this index structure achieves a good performance, for each 12-mer from reads, it is only needed an array access to find out the start position of its mapped intervals in the $hv\_pos$, and it is very friendly to the cache of MP since these intervals are stored in a continuous way, and easily coped to buffers for SPs to fetch.
% \begin{figure}[!htb]
% 	\begin{center}
% 		\includegraphics[width=0.9\linewidth]{HIdx}
% 		\caption{Memory layout of the hash index. For the sake of simplicity, we use 3-mers instead of 12-mers.}
% 		\label{HIdx}
% 	\end{center}
% \end{figure}

% \subsubsection{Mixed Sequence Packing}
% {\color{green} Maybe move this to SIMD section}
% In order to efficiently compute the contributions of the substitution term $eq_{i,j}$ (see Section \ref{SecMyers}) we transform both the read and reference into a bit-vector representation. The literature states three major packing strategies: continuous packing, separated packing and one-hot coding. The first one simply stores the sequences in a single bit-vector where the even bits encode the lower and the odd bits the upper part of the 2 bit representation of nucleotides, so to speak, as array of structs (like BWA). The second strategy employs two bit-vectors to separate the even and odd contributions (struct of arrays). The third strategy interprets the four base pairs as class labels and subsequently stores their characteristic functions in four bit-vectors (like RazerS3).

% %There are mainly two types of packing strategy, continuous packing and separated packing, continuous packing strategy packs sequence to a bit vector, and each bp in the sequence is implied by two adjacent bits in the vector (like BWA), while separated packing packs sequence to two bit vectors, one for higher bits and the other for lower bits or use just 4 bit-vectors to indicate whether a bp is 'A', 'C', 'G', or 'T' (like RazerS3).

% %Both of the two packing strategy have their advantages, the continuous one have better cache performance while the separated one can gain better performance while computing. So we used a mixed packing strategy. We pack every 64 bps in the sequence using the separated packing strategy, thus we get two 64-bit integers, and put them in two adjacent position in the packed data array, which is not only friendly to computing but also friendly to cache. Fig.\ref{MixPack} shows differences among these packing strategy.

% Continuous packing is very cache-friendly while the latter two provide better performance during computation. Our approach uses a mixed packing strategy. We pack 64 contiguous base pairs using the separated packing strategy, resulting in two 64 bit integers, and subsequently put them in two adjacent positions of the array. The resulting memory layout is both cache-friendly and SIMD-compliant (see Figure~\ref{MixPack}).

%Since there are over 13,000 nodes involved in our experiment with over 50,000 processes dispatched, $\sim$800GB of reads data processed in the mass parallel evaluation, simply dividing reads data to 50,000 pieces with a size of $\sim$16MB is obviously not a good solution.

%Thus, we use task grid pattern for parallelization among CGs, as we have divided references to blocks and reads to chunks, each chunk of read should be mapped against each reference block, we take a comparison between a chunk of read and a reference block as an independent task, each task can be assigned to any process and each process can process arbitrary number of tasks. 
%Then we organize tasks to a task grid, and let each task in the same row share the same reference block, while each task in the same column share the same read data. Each process processes several cells in a row, which is the comparison between some read chunks and a reference block. It is because that the hash index of references are relatively large, we do not want to reload another one unless the number of processes is less than the number of reference blocks. And in the case processes count fewer than reference blocks count, we can just process the grid rows by rows.

%In Sunway Taihu Light's file system, each node are connected to the storage with a 1GBit/s network, so the total bandwidth should be $40,000 * 1GBit/s = 40TBit/s$, but it is obviously the total bandwidth of its storage can not reach such a high bandwidth, and conflicts may happen if there are too many file system requests which can reduce the bandwidth of data loading. In our experiments, we failed to wait processes finishing loading their data if each process loads its reference and reads by itself.

%Also, because there are 4 CGs in one node with shared file system bandwidth, so there is no good to let them to do input/output at the same moment, making it problematic to assign processes to the task grid.

%then processes in the first row or column of the task grid load their reads and references, then data loaded by these processes is broadcasted to their row group or column group.

>>>>

\documentclass[conference]{IEEEtran}

\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage{cases}
\usepackage{listings}
\lstset{language=C++,
    basicstyle=\ttfamily \footnotesize,
    keywordstyle=\color{black}\ttfamily,
    stringstyle=\color{black}\ttfamily,
    commentstyle=\color{gray}\ttfamily,
    morecomment=[s][\color{blue}]{/**}{**/},
    showstringspaces=false,
    showtabs=false,
    captionpos=b,
    numbers=left
}

\hyphenation{cluster computational-biology short-read-mapping}

@@

\title{S-Aligner: Ultra-scalable read mapping on Sunway Taihu Light}

\author{\IEEEauthorblockN{Xiaohui Duan, Kai Xu, Yuandong Chan}
\IEEEauthorblockA{%College of Computer Science and Technology \\
Shandong University\\
Jinan, China\\
Email: \{sunrise.duan, xu\_kai, yuandong.c\}@mail.sdu.edu.cn}
\and
\IEEEauthorblockN{Christian Hundt, Bertil Schmidt}
\IEEEauthorblockA{Johannes Gutenberg University\\
Mainz, Germany\\
Email: \{hundt, bertil.schmidt\}@uni-mainz.de}
\and
\IEEEauthorblockN{Pavan Balaji}
\IEEEauthorblockA{Argonne National Laboratory\\
Argonne, Illinois, USA\\
Email: balaji@anl.gov}
\and
\IEEEauthorblockN{Weiguo Liu}
\IEEEauthorblockA{Shandong University\\
Jinan, China\\
Email: weiguo.liu@sdu.edu.cn\\
}
}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=\linewidth]{figures/SW26010}
		\caption{Architecture of the SW26010 processor, which is partitioned into four CGs.}
		\label{fig:SW26010}
	\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\includegraphics[width=1\linewidth]{figures/BitDP}
	\caption{Relations between the original matrix, $\varDelta v$, $V^+$, and $V^-$.  Here (a) shows the original score matrix; light-shaded cells indicate $\varDelta v_{i,j}=+1$ while heavily shaded cells correspond to $\varDelta v_{i,j}=-1$. The vertical derivative $\varDelta v$ shown in (b) is subsequently one-hot encoded by the bit vectors $V^+$ and $V^-$ in (c) and (d), respectively.}
	\label{BitDP}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{figures/GridNew}
	\caption{Our task-partitioning and file-loading strategy: (a) overall design; (b) and (c) detailed behavior of processes.}
	\label{TaskGrid}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=0.9\linewidth]{figures/RefDiv}
		\caption{Exemplary partitioning of a reference genome based on concatenation and padded splits.}
		\label{RefDiv}
	\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{figures/AsyncRead}
	\caption{Asynchronous data loading strategy: (a) default case, and (b) case when there is no significantly shorter reference block.}
	\label{AsyncRead}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{figures/FrmWk}
	\caption{Workflow of S-Aligner on a single CG: gray boxes correspond to tasks assigned to SPs while tasks in white boxes are executed on the MP.}
	\label{FrmWk}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{figures/AsyncPara}
	\caption{Asynchronous filtration.}
	\label{AsyncPara}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=\linewidth]{figures/SPParallel}
	\caption{Framework of our parallelization scheme within one CG: the MP performs filtration and output of alignment results while SPs verify identified candidate locations and generate base-pair level alignments. Gray boxes indicate the intervals verified in the second round of verification (assuming $slice\_num = 4$).}
	\label{SPParallel}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=0.9\linewidth]{figures/MixPack}
	\caption{Bit-level encoding strategies for DNA sequences. For simplicity, we use a 4-bit word instead of 64-bit word for illustrating our combined AoSoA strategy (bottom).}
	\label{MixPack}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	%\begin{center}
	\includegraphics[width=1\linewidth]{figures/AsyncTrans}
	\caption{Asynchronous data transfer to LDM.}
	\label{AsyncTrans}
	%\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/VarVerCha}
		\caption{Comparison of the runtimes of four implementations of Myers' algorithm on a single CG.}
		\label{VarVerCha}
	\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/VarOptCha}
		\caption{Evaluation with and without I/O optimizations.}
		\label{VarOptCha}
	\end{center}
\end{figure}

@@

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/VarProcCha}
		\caption{Speedup and efficiency of S-Aligner for different numbers of nodes.}
		\label{VarProcCha}
	\end{center}
\end{figure}

@@

\bibliographystyle{plain}
\bibliography{bib/references.bib}

% that's all folks
\end{document}
